Investigating the Gender and Nationality Distributions of the MoMA: While Modern, Not Diverse in Its Artists and Decision-making Staff

### 2023-04-19 outputs/paper/paper.qmd
Wrote a draft of the analysis paper

---
title: "How Diverse is the MoMA: While Modern, Not Diverse in Its Artists and Directors"
subtitle: "An analysis of the MoMA's public datasets on their collections and exhibits"
author: Sakura Ariga
thanks: "Code and data are available at: https://github.com/sakura-ariga/MOMA_diversity_analysis.git."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "The MoMA's public GitHub data on their exhibits and collections was analyzed to determine how diverse the artists and directors at the MoMA are. It was found that the MoMA is comprised largely of white men, but that this relationship has changed with time. This finding gives quantitative proof of the need for more representation within not jus the MoMA, but all American museums as a whole."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

It was found that the MoMA's artists and directors were predominantly men. Simple linear regression models show that time affects this percentage of female artists, directors, and department heads in that the number of female artists has increased on average by ___ percentage points per year, while the number of female directors and department heads has increased on average by 0.25 percentage points per year. These findings are important because they provide numerical proof of the lack of gender diversity in the MoMA, both in the artists displayed and in the directors and department heads who choose the artists to display, which could motivate the MoMA to take active steps to better the gender representation within their institution.

@@

@tbl-exhibits shows the 7 variables of the MoMA exhibition index dataset and the first 10 rows of it, where each row represents a different individual constituent in the exhibit. The Exhibition Title variable refers to the exhibit name. The Name variable indicates the artist. Nationality indicates the individual's nationality (e.g. "American") and is investigated further in @_____. Gender has two options of "Male" or "Female". Exhibition Start and Exhibition End indicate the time period that the exhibit was displayed at the MoMA for, and are the only numerical variables in this dataset.

```{r}
#| echo: false
#| eval: true
#| label: tbl-exhibits
#| tbl-cap: "First ten rows of the MoMA Exhibits dataset"
#| warning: false

cleaned_momaexhibit_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Exhibition Title", "Artist Name", "Nationality", "Gender", "Exhibition Start", "Exhibition End"),
    booktabs = TRUE,
    linesep = ""
  )
```

@tbl-directors shows the 6 variables of the MoMA directors and department heads dataset and the first 10 rows of it, where each row represents a different individual. The Department variable refers to the specific curatorial department within the MoMA that the individual is in charge of. As with the exhibition dataset, the Name variable indicates the person, Nationality indicates the individual's nationality (e.g. "American") and is investigated further in @_____, and Gender has two options of "Male" or "Female". Start and End indicate the time period that the individual held the position for, and are the only numerical variables in this dataset.

@@

cleaned_momadirectors_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Department", "Name", "Start", "End", "Nationality", "Gender"),
    booktabs = TRUE,
    linesep = ""
  )
```

@@

@fig-artistnationality displays the nationality distribution of the artists whose works have been displayed in exhibits in the MoMA. Aside from the large number of missing values (marked as NA at the rightmost bar), the most common nationality among is American, at 14,112 artists. The second highest nationality is French, at 3,747 artists - nearly a fifth of the American artists. The large number of American artists may be explained in part because the MoMA itself is American. It is surprising though that while the MoMA has presented 2,500+ exhibits throughout its history, \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_.

@@

# Bar graph showing nationality distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  geom_bar(position = "dodge2")+
  labs(x = "Nationality", y = "Number of Artists in MoMA Exhibits") 
```

@@

@fig-artistgender displays the gender distribution of the artists whose works have been displayed in exhibits in the MoMA. As the graph demonstrates, there is a much larger number of male artists who have been featured in the MoMA exhibits than female artists. With 22,518 male artists and 2,301 female artists, the male to female ratio is approximately 10:1, showing significant inequality in gender representation in MoMA artists. This may be partly explained by _____.

@@

# Bar graph showing gender distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |> 
  count(Gender) |>
  ggplot(mapping = aes(x = Gender, y = n)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(x = "Gender", y = "Count of Artists in MoMA Exhibits") 
```

@@

@fig-staffnationality displays the nationality distribution of all of the directors and department heads at the MoMA. Here too, like in the nationality distribution of artists, there is a majority of American nationality among the directors and department heads, at 50 Americans. Compared to the 3 Swiss, 3 British, 2 German, 1 French, and 1 Austrian directors and department heads, \_\_\_\_\_\_\_\_\_\_.

@@

# Bar graph showing nationality distribution of MoMA directors and department heads 
cleaned_momadirectors_data |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  geom_bar(position = "dodge2")+
  labs(x = "Nationality", y = "Count of MoMA Directors and Department Heads") 
```

@@

@fig-staffgender displays the gender distribution of all of the directors and department heads at the MoMA. Here too, there are many more male directors and department heads compared to female department heads. With 53 male directors and department heads and 12 female directors and department heads, the male to female ratio is approximately 4:1, showing inequality in gender representation in MoMA leadership. This may be partly explained by \_\_\_\_\_\_\_\_\_\_.

@@

# Bar graph showing gender distribution of MoMA directors and department heads 
cleaned_momadirectors_data |>
  ggplot(mapping = aes(x = Gender)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  geom_bar(position = "dodge2")+
  labs(x = "Gender", y = "Count of MoMA Directors and Department Heads")
```

@@

### Model Justification

@@

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| label: fig-model1
#| fig-cap: Linear regression model of the effect of time on the percentage of female MoMA directors
#| echo: false
#| eval: true
#| warning: false
#| message: false

model_directors_percentage_female <-
  readRDS(file = here::here("outputs/models/model_directors_percentage_female.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "First model" = model_directors_percentage_female
  ),
  fmt = 2
)
```

```{r}
#| label: fig-model2
#| fig-cap: Linear regression model of the effect of exhibit year on the percentage of female MoMA artists
#| echo: false
#| eval: true
#| warning: false
#| message: false

model_exhibits_percentage_female <-
  readRDS(file = here::here("outputs/models/model_exhibits_percentage_female.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "Second model" = model_exhibits_percentage_female
  ),
  fmt = 2
)
```

```{r}
#| label: fig-model3
#| fig-cap: Linear regression model of the effect of time and percentage of female MoMA directors on the percentage of female MoMA artists selected for exhibits
#| echo: false
#| eval: true
#| warning: false
#| message: false

model_exhibits_on_directors <-
  readRDS(file = here::here("outputs/models/model_exhibits_on_directors.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "Third model" = model_exhibits_on_directors
  ),
  fmt = 2
)
```

@@

## Gender distribition {#sec-first-point}

## Nationality distribition {#sec-second-point}

## Analyzing the model: is time a factor? {#sec-third-point}

## Comparing artists vs directors {#sec-fourth-point}

## Weaknesses and next steps

-   a LOT of NAs in the exhibit dataset, especially regarding artist gender, so results are skewed bc of that
-   binary understanding of gender that is exclusive

>>>>

---
title: "Investigating the Gender and Nationality Distributions of the MoMA: While Modern, Not Diverse in Its Artists and Decision-making Staff"
subtitle: "An analysis of the MoMA's public datasets"
author: Sakura Ariga
thanks: "Code and data are available at: https://github.com/sakura-ariga/MOMA_diversity_analysis.git."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "The MoMA's public GitHub data on their exhibits and collections was analyzed to determine how diverse the artists and directors at the MoMA are in terms of gender and nationality. It was found that the MoMA is comprised largely of American men and particularly lacking in their representation of women, both as artists and as directors and department heads. However, it was found that women representation has a positive relationship with time, indicating that the MoMA's gender distribution has improved slightly over time. This finding gives quantitative proof of the need for more gender representation in the MoMA, both in the artists displayed and in the top decision-making staff hired."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

It was found that the MoMA's artists and directors were predominantly men. Simple linear regression models show that time affects this percentage of female artists, directors, and department heads in that the number of female artists has increased on average by 0.08 percentage points per year, while the number of female directors and department heads has increased on average by 0.24 percentage points per year. It was also found that the percentage of female directors and department heads has a positive relationship with the percentage of female artists displayed, with an increase of an average of 0.10 percentage points in women artists with every 1% increase of women directors and department heads. These findings are important because they provide numerical proof of the lack of gender diversity in the MoMA, both in the artists displayed and in the directors and department heads who choose the artists to display, which could motivate the MoMA to take active steps to better the gender representation within their institution.

@@

@tbl-exhibits shows the 7 variables of the MoMA exhibition index dataset and the first 10 rows of it, where each row represents a different individual constituent in the exhibit. The Exhibition Title variable refers to the exhibit name. The Name variable indicates the artist. Nationality indicates the individual's nationality (e.g. "American") and Gender has two options of "Male" or "Female". Exhibition Start and Exhibition End indicate the time period that the exhibit was displayed at the MoMA for, and are the only numerical variables in this dataset.
\newpage
```{r}
#| echo: false
#| eval: true
#| label: tbl-exhibits
#| tbl-cap: "First ten rows of the MoMA Exhibits dataset"
#| warning: false

# Create table of first ten rows of the cleaned_momaexhibit_data to explain variables 
cleaned_momaexhibit_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Exhibition Title", "Artist Name", "Nationality", "Gender", "Exhibition Start", "Exhibition End"),
    booktabs = TRUE,
    linesep = ""
  ) |>
  kable_styling(latex_options="scale_down")
```

@tbl-directors shows the 6 variables of the MoMA directors and department heads dataset and the first 10 rows of it, where each row represents a different individual. The Department variable refers to the specific curatorial department within the MoMA that the individual is in charge of. As with the exhibition dataset, the Name variable indicates the person, Nationality indicates the individual's nationality (e.g. "American"), and Gender has two options of "Male" or "Female". Start and End indicate the time period that the individual held the position for, and are the only numerical variables in this dataset.

@@

# Create table of first ten rows of the cleaned_momadirectors_data to explain variables 
cleaned_momadirectors_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Department", "Name", "Start", "End", "Nationality", "Gender"),
    booktabs = TRUE,
    linesep = ""
  ) |>
  kable_styling(latex_options="scale_down")
```

@@

@fig-artistnationality displays the nationality distribution of the artists whose works have been displayed in exhibits in the MoMA. Aside from the large number of missing values (marked as NA at the rightmost bar), the most common nationality among is American, at 14,112 artists. The second highest nationality is French, at 3,747 artists - nearly a fifth of the American artists. The large number of American artists may be explained in part because the MoMA itself is American.

@@

# Bar graph showing nationality distribution of artists displayed in MoMA exhibits
cleaned_momaexhibit_data |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1, size = 5)) +
  geom_bar(position = "dodge2", fill = "cornflowerblue")+
  labs(x = "Nationality", y = "Number of Artists in MoMA Exhibits") 
```

@@

@fig-artistgender displays the gender distribution of the artists whose works have been displayed in exhibits in the MoMA. As the graph demonstrates, there is a much larger number of male artists who have been featured in the MoMA exhibits than female artists. With 22,518 male artists and 2,301 female artists, the male to female ratio is approximately 10:1, showing significant inequality in gender representation in MoMA artists.

@@

# Bar graph showing gender distribution of artists displayed in MoMA exhibits
cleaned_momaexhibit_data |> 
  count(Gender) |>
  ggplot(mapping = aes(x = Gender, y = n)) +
  geom_col(fill = "mediumorchid") + 
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(x = "Gender", y = "Count of Artists in MoMA Exhibits") 
```

@@

@fig-staffnationality displays the nationality distribution of all of the directors and department heads at the MoMA. Here too, like in the nationality distribution of artists, there is a majority of American nationality among the directors and department heads, at 50 out of 60 Americans.

@@

# Bar graph showing nationality distribution of MoMA directors and department heads 
cleaned_momadirectors_data |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  geom_bar(position = "dodge2", fill = "cornflowerblue")+
  labs(x = "Nationality", y = "Count of MoMA Directors and Department Heads") 
```

@@

@fig-staffgender displays the gender distribution of all of the directors and department heads at the MoMA. Here too, there are many more male directors and department heads compared to female department heads. With 53 male directors and department heads and 12 female directors and department heads, the male to female ratio is approximately 4:1, showing inequality in gender representation in MoMA leadership.

@@

# Bar graph showing gender distribution of MoMA directors and department heads 
cleaned_momadirectors_data |>
  ggplot(mapping = aes(x = Gender)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  geom_bar(position = "dodge2", fill = "mediumorchid")+
  labs(x = "Gender", y = "Count of MoMA Directors and Department Heads")
```

@@

## Model Justification

@@

# Results
The results of the first linear regression model of the effect of time on the percentage of female MoMA directors is summarized in @tbl-modelresults1. The prediction of a positive relationship between time in years and the percentage of female MoMA directors proved correct, with the percentage of female directors increasing every year by 0.25 percentage points since the MoMA's founding in 1929. We know that this finding is significant because the p-value of this model is less than the threshold of 0.05. @fig-model1 demonstrates this  relationship as a scatterplot, better visualizing the positive linear relationship between the amount of women leaders in the MoMA and time.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Import linear regression model of the effect of time on percentage of female MoMA directors
model_directors_percentage_female <-
  readRDS(file = here::here("outputs/models/model_directors_percentage_female.rds"))
```

@@

# Table showing results of model of the effect of time on percentage of female MoMA directors
modelsummary::modelsummary(
  list(
    "First model" = model_directors_percentage_female
  ),
  fmt = 2
)
```

```{r}
#| label: fig-model1
#| fig-cap: Linear model of percentage of female MoMA directors by year
#| echo: false
#| warning: false

model_directors_percentage_female |>
  ggplot(aes(x = Year, y = PercentageFemale)) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  ) +
  labs(
    x = "Time (Years)",
    y = "Percentage of Female MoMA Directors"
  ) 
```
\newpage
The results of the second linear regression model of the effect of time on the percentage of female MoMA artists is summarized in @tbl-modelresults2. The prediction of a positive relationship between time in years and the percentage of female MoMA artists proved correct, with the percentage of female artists increasing every year by 0.08 percentage points since the MoMA's founding in 1929. This finding is significant because the p-value of this model is less than the threshold of 0.05. @fig-model2 demonstrates this  relationship as a scatterplot, better visualizing the positive linear relationship between the amount of women artists who have been displayed in MoMA exhibits and time. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Import linear regression model of the effect of time on percentage of female MoMA artists
model_exhibits_percentage_female <-
  readRDS(file = here::here("outputs/models/model_exhibits_percentage_female.rds"))
```

@@

# Table showing results of model of the effect of time on percentage of female MoMA artists
modelsummary::modelsummary(
  list(
    "Second model" = model_exhibits_percentage_female
  ),
  fmt = 2
)
```

```{r}
#| label: fig-model2
#| fig-cap: Linear model of percentage of female MoMA artists by year
#| echo: false
#| warning: false

model_exhibits_percentage_female |>
  ggplot(aes(x = Year, y = PercentageFemale)) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "black",
    linetype = "dashed",
    formula = "y ~ x"
  ) +
  labs(
    x = "Time (Years)",
    y = "Percentage of Female MoMA Artists"
  ) 
```

The results of the third linear regression model of the effect of time and the percentage of female MoMA directors on the percentage of female MoMA artists is summarized in @tbl-modelresults3. The prediction of a positive relationship between time in years and the percentage of female directors and the percentage of female MoMA artists proved correct, with the percentage of female artists increasing by 0.10 percentage points with every 1% increase in the percentage of female directors. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Import linear regression model of the effect of time and percentage of female MoMA directors on percentage of female MoMA artists
model_exhibits_on_directors <-
  readRDS(file = here::here("outputs/models/model_exhibits_on_directors.rds"))
```

@@

# Table showing results of model of the effect of time and percentage of female MoMA directors on percentage of female MoMA artists
modelsummary::modelsummary(
  list(
    "Third model" = model_exhibits_on_directors
  ),
  fmt = 2
)
```

@@

## Gender Distribution 
From @fig-artistgender and @fig-staffgender, we see that there is a significant gap between representation of men versus women in the MoMA, both for the artists displayed and for the museum staff with the most power. Comparing @fig-artistgender and @fig-staffgender, the men:women ratio is more unequal in the artists, at a ratio of 10:1, than in the directors and department heads, at a ratio of 4:1.

This distribution is surprising, especially given that the MoMA was founded due to the works and ideas of women: Lillie P. Bliss, Mary Quinn Sullivan, and Abby Aldrich Rockefeller - three influential women who were patrons of the arts in the 1920s [@MoMA]. One potential reason for this could be American society's historical attitude towards women versus men, tending to elevate and encourage men to pursue careers as artists or museum curators and discouraging women to. Another reason could be related to the findings of the third model found in @tbl-modelresults3: if women directors and department heads are more likely to choose women artists' work to exhibit, the same may go for men directors and department heads with men artists. These two reasons could be related to each other as well, with male directors and department heads more likely to choose male artists due to social conditioning from society's expectations of gender roles.

While the above models (found in @fig-model1, @fig-model2, @tbl-modelresults3) show that change has occurred and that the percentage of women represented in the MoMA's displayed artists and decision-making staff has increased over time, there is still much more room for improvement. As an institution with "modern" in its name, the MoMA should act as a role model for other cultural institutions and be an active leader in incorporating more diversity in the museum, both in showcasing more female artists and in hiring more female directors and department heads. This holds true especially now that it has been nearly a century since the MoMA's opening and thus much more modern art has been created and can be displayed.

## Nationality Distribution 
@fig-artistnationality and @fig-staffnationality show us that the museum is predominantly made up of American representation, both in its artists and in its decision-making staff. Most interesting is that the nationalities most represented in the artists is also represented well in the directors and department heads: American, French, British. One possible explanation of this is that these three countries place a heavy emphasis on the modern arts, and so produce many artists and museum staff. Another explanation could be that the MoMA places a eurocentric, Western-heavy emphasis on their exhibits, and thus is biased towards selecting American, French, and British artists and staff. Because nationality is not the best indicator for social inclusion practices, analysis on racial discrimination cannot be made here from this data.

## Weaknesses and Next Steps
One weakness of this analysis is the binary understanding of gender as only male or female. This excludes artists, directors, and department heads who do not identify within this gender binary and thus this analysis was not able to completely measure gender diversity in the MoMA. Another bias present in this analysis is the removal of missing values that occurred during the data cleaning step of this paper, which made up a large number of responses, particularly in the gender of artists. This inevitably led to the results of the gender distribution analysis being biased and not entirely accurate.

Thus, an appropriate next step would be to conduct more research on the demographics of the artists, directors, and department heads of the MoMA so that a more accurate and meaningful analysis can be conducted. Not only would this account for some of the missing values that were excluded in this paper, this could result in analysis on racial equality in the MoMA. While the nationality variable that was analyzed in this paper was helpful in understanding some aspect of inclusion in the MoMA, in a museum that is so dominated by Americans, both as artists and as decision-making museum staff, an analysis on the racial representation in the MoMA would yield more actionable insights.
